### 2020-04-07 paper/commands.tex
section 2

    % commands
    \renewcommand\thesubfigure{(\alph{subfigure})}
    \newcommand{\tDelta}{\textsf{Delta}\xspace}

    % tikz
    \usetikzlibrary{decorations.pathmorphing}
    \tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
    \tikzstyle{big_vertex} = [circle, minimum width=40pt,draw,inner sep=0pt]
    \tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
    \tikzstyle{local} = [thick,->,>=stealth]
    \tikzstyle{dist} = [thick,dashed,->,>=stealth]
    \tikzstyle{node}=[circle,draw,radius=0.3, align=center]

    % colours
    \definecolor{grey}{rgb}{0.52, 0.52, 0.51}
    \definecolor{red}{rgb}{0.7, 0.11, 0.11}
    \definecolor{blue}{rgb}{0.0, 0.0, 0.55}
    \definecolor{green}{rgb}{0.0, 0.42, 0.24}

            \def\record(#1,#2,#3,#4,#5,#6){
              \node (v) [draw,black,minimum height=0.7cm,minimum width=1cm,text width=0.2\linewidth,xshift=-6cm,align=center,yshift=#1cm] {#2};

            \node (b) [draw,black,minimum height=0.7cm,minimum width=11.4cm,text width=0.2\linewidth,xshift=1.4cm,yshift=#1cm] {};

            \draw[thick,->] (v) -- (b);

            \node (i1)  [draw,black,minimum height=0.5cm,minimum width=2cm,xshift=-2.8cm,text width=0.3\linewidth,yshift=#1cm,align=center] {#3};

            \node (i2)  [draw,black,minimum height=0.5cm,minimum width=2.5cm,xshift=0.9cm,right of=i1,yshift=0cm,text width=0.3\linewidth,align=center] {#4};

            \node (i3)  [draw,black,minimum height=0.5cm,minimum width=2.5cm,xshift=0.75cm,right of=i2,yshift=0cm,align=center] {#5};

            \node (i4)  [draw,black,minimum height=0.5cm,minimum width=2.5cm,xshift=0.6cm,right of=i3,yshift=0cm,align=center] {#6};

        }

\def\spaceRecord(#1){
\node (v) [black,minimum height=0.7cm,minimum width=1cm,text width=0.2\linewidth,xshift=-6cm,align=center,yshift=#1cm] {$\vdots$};
\node (b) [black,minimum height=0.7cm,minimum width=11.4cm,text width=0.2\linewidth,xshift=1.4cm,yshift=#1cm] {};
\node (i1)  [black,minimum height=0.5cm,minimum width=2cm,xshift=-2.8cm,text width=0.3\linewidth,yshift=#1cm,align=center] {$\vdots$};
\node (i2)  [black,minimum height=0.5cm,minimum width=2.5cm,xshift=0.9cm,right of=i1,yshift=0cm,text width=0.3\linewidth,align=center] {$\vdots$};
\node (i3)  [minimum height=0.5cm,minimum width=2.5cm,xshift=0.75cm,right of=i2,yshift=0cm,align=center] {$\vdots$};
\node (i4)  [minimum height=0.5cm,minimum width=2.5cm,xshift=0.6cm,right of=i3,yshift=0cm,align=center] {$\vdots$};
}