### 2020-04-10 paper/sec2-reciprocal-consistency.tex

\section{Reciprocal Consistency}
\label{sec:recipr-cons}

In the property graph data model, edges have direction and each edge runs from a \emph{source} vertex to a \emph{destination} vertex.
In the storage layer, however, edge directionality does not exist;  \textbf{both} the source and the destination vertices store information about each other.
This allows edge traversal to be bidirectional and speeds up query performance.

Consider, for example, the statement: ``Tolkien \textit{wrote} The Hobbit''.
It is expressed using vertex  \emph{a} for Tolkien and vertex \emph{b} for The Hobbit, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination). Corresponding openCypher \cite{openCypher} code is given below and Fig \ref{rc-edge} shows the model level view.% of edge $ab$.
\input{cypher/create-edge}

Fig \ref{db-rep} depicts the internal representation of a graph arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}.
A vertex, such as \emph{a}, is represented by a record that contains one or more properties of that vertex, followed by a sequence of \emph{edge pointers} pointing to all those vertices to which this vertex is related either as a source or a destination.
The sequence of edge pointers is also called the \emph{adjacency list}.

It can be seen in Fig \ref{db-rep} that \emph{a}'s adjacency list has an edge pointer entry that stores `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s list has a corresponding entry storing the reciprocal (or inverse) information `\emph{b} \emph{written} by \emph{a}'.
When the adjacency list entries for a given edge refer to each other in a complementary manner like this, that edge is said to be \emph{reciprocally consistent}.

Consider a query: `list all titles by the author who wrote The Hobbit'.
This query needs to start from \emph{b} which represents the only entity specified explicitly in it.
Thanks to the reciprocal information in \emph{b}, it can reach \emph{a} from \emph{b}, even though edge \emph{ab} is  ``directed'' from \emph{a} to \emph{b}, and then compile the necessary list from \emph{a}.
Note that reciprocal consistency is assumed to be prevail when a query reads only the source or destination vertex of an edge.

\input{figures/rc}