### 2020-04-16 slides/figures/graph.tex
copy epew slides as base

\begin{figure}[h!]
  \centering
  \resizebox{0.6\textwidth}{!}{
    \begin{tikzpicture}[node distance=2cm]
      \node (v1) [vertex] {};
      \node (v2) [vertex, above of=v1,xshift=0.3cm] {};
      \node (v3) [vertex, above right of=v1,yshift=-0.3cm] {};
      \node (v4) [vertex, right of=v3,yshift=-0.3cm] {};
      \node (v5) [vertex, above left of=v4,xshift=0.7cm] {};
      \node (v6) [vertex, right of=v5,xshift=-0.8cm,yshift=1.2cm] {};
      \node (v7) [vertex, below right of=v6, yshift=-0.8cm] {};
      \node (v8) [vertex, below of=v7, yshift=0.3cm,xshift=0.3cm] {};
      \node (v9) [vertex, right of=v7,xshift=-0.4cm,yshift=0cm] {};
      \node (v10) [vertex, above left of=v9,xshift=0.4cm,yshift=-0.4cm] {};
      \node (v11) [vertex, above right of=v10,yshift=-2.5cm,xshift=1cm] {};
      \node (v12) [vertex, above of=v10,yshift=-0.6cm,xshift=1cm] {};
      \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
      \draw [local] (v2) -- (v1);
      \draw [local] (v1) -- (v3);
      \draw [local] (v3) -- (v2);
      \draw [local] (v2) -- (v5);
      \draw [local] (v1) -- (v4);
      \draw [local] (v5) -- (v4);
      \draw [local] (v5) -- (v6);
      \draw [local] (v4) -- (v7);
      \draw [local] (v8) -- (v4);
      \draw [local] (v8) -- (v9);
      \draw [local] (v9) -- (v11);
      \draw [local] (v10) -- (v7);
      \draw [local] (v10) -- (v12);
      \draw [local] (v12) -- (v13);
      \draw [local] (v10) -- (v6);
      \draw [local] (v10) -- (v11);
      \draw [local] (v13) -- (v10);
      \draw [local] (v11) -- (v12);
    \end{tikzpicture}
  }
  \caption{Connected graph}
\end{figure}