### 2020-04-20 slides/figures/rc-edge.tex

\begin{tikzpicture}[node distance=2.2cm]
  \node (v1) [big_vertex,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
  \node (v2) [big_vertex,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
  \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
  \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
  \draw [thick,->,>=stealth] (0.9,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} (4.1,0);
\end{tikzpicture}