### 2020-04-20 slides/main.tex

\documentclass[aspectratio=169]{beamer}

\input{config}

\input{title}

\begin{document}

\input{title-images}

\setcounter{framenumber}{0}
\metroset{numbering=counter}

\input{s1-property-graph}

\input{s2-storage-layer}

\input{s3-reciprocal-consistency}

\input{s4-dist-graph}

\input{s6-half-corruption}

\input{s7-delta}

\input{s8-evaluation}

\input{s9-summary}

% %------------------------------------------------
% % Problem 1: Reciprocal Consistency}
% % ------------------------------------------------

% \begin{frame}
%   \frametitle{Reciprocal Consistency Violation}
%   Two concurrent transactions:
%   \begin{itemize}
%   \item Mr Red requests to book the seat, $T_R$
%   \item Mr Blue requests to book the seat, $T_B$
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Reciprocal Consistency Violation}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}[node distance=2cm]
%       \uncover<1-> {
%         \draw [black,<-,>=stealth] (0,0) -- (0,4) node [] {};
%         \draw [black,<-,>=stealth] (4,0) -- (4,4) node [] {};
%         \draw [black,dashed,->,>=stealth] (0,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (4,4.1) node [label=above:{$S_2$}] {};
%       }
%       \uncover<2-3> {
%         \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (2,2.3);
%       }
%       \uncover<3-4> {
%         \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (2,2.1);
%       }
%       \uncover<4-> {
%         \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (4,1);
%       }
%       \uncover<5-> {
%         \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (0,1.2);
%       }
%     \end{tikzpicture}
%     \caption{Concurrent transactions interleave and violate reciprocal consistency}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Reciprocal Consistency Violation}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}[node distance=7cm]
%       \node (n1) [bignode] {\texttt{Flight}};
%       \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
%       \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
%       \node (n2) [bignode, right of=n1] {\texttt{Seat}};
%       \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);

%       \uncover<2-> {
%       \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Mr Blue Booked In Seat}};
%       \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Mr Red Booked In Seat}};
%       \draw [<->,dashed](q1) -- (q2);
%       \draw [<->,dashed](q1) -- (q2);
%       \coordinate (a) at (4,-2);
%       \coordinate (b) at (3,-3);
%       \coordinate (c) at (3,-2);
%       \coordinate (d) at (4,-3);
%       \draw [very thick,red] ($ (a) $) -- ($ (b) $);
%       \draw [very thick,red] ($ (c) $) -- ($ (d) $);
%     }
%     \end{tikzpicture}
%     \caption{Reciprocal inconsistent distributed edge}
%   \end{figure}
% \end{frame}

% % \begin{frame}
% %   \frametitle{Reciprocal Consistency Violation}
% %   \begin{figure}[h!]
% %     \centering
% %     \begin{tikzpicture}[node distance=2cm]
% %       \uncover<1->{
% %         \draw [black,<-,>=stealth] (4.5,0) -- (4.5,4) node [] {};
% %         \draw [red,->,>=stealth] (4.5,2.2) node [label=left:{$T_x$}] {} -- (6.5,1);
% %         \draw [blue,->,>=stealth] (6.5,2.7) node [label=right:{$T_y$}] {} -- (4.5,1.2);
% %         \draw [black,<-,>=stealth] (6.5,0) -- (6.5,4) node [] {};
% %         \draw [black,dashed,->,>=stealth] (4.5,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (6.5,4.1) node [label=above:{$S_2$}] {};
% %       }
% %       \uncover<2->{
% %         \draw [black,<-,>=stealth] (9,0) -- (9,4) node [] {};
% %         \draw [red,->,>=stealth] (9,3.5) node [label=left:{$T_x$}] {} -- (11,1);
% %         \draw [blue,->,>=stealth] (9,2.8) node [label=left:{$T_y$}] {} -- (11,2);
% %         \draw [black,<-,>=stealth] (11,0) -- (11,4) node [] {};
% %         \draw [black,dashed,->,>=stealth] (9,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (11,4.1) node [label=above:{$S_2$}] {};
% %       }
% %     \end{tikzpicture}
% %     \caption{Possible interleaving of concurrent transactions that violate reciprocal consistency}
% %   \end{figure}
% % \end{frame}

% \begin{frame}
%   \frametitle{Distributed Edge Reciprocal Consistency}
%   \begin{itemize}
%     \uncover<1->{
%   \item Without concurrency control, distributed edges can become reciprocally inconsistent
%     \begin{itemize}
%     \item A distributed edge's reciprocal information is separated by the network
%     \end{itemize}
%   }
%   \uncover<2->{
%   \item Reciprocal inconsistency is a source of corruption\footnotemark
%   }
%   \uncover<3->{
%   \item No known concurrency control protocol exists specific to graph databases and maintaining reciprocal consistency
%     }
%     \end{itemize}
%     \footnotetext[3]{Ezhilchelvan, P. et al, \textit{On the degradation of distributed graph databases with eventual consistency}. European Performance Engineering Workshop 2018.}
% \end{frame}

% \begin{frame}
%   \frametitle{Our Solution: Collision Detection}
%   \begin{itemize}
%     \uncover<1->{
%     \item Carefully abort transaction(s)
%     }
%     \uncover<2->{
%     \item No centralized control or synchronized clock
%     }
%     \uncover<3->{
%     \item Permits interleavings that preserve reciprocal consistency
%     }
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Collision Detection General Rules}
%   \begin{itemize}
%     \uncover<1->{
%     \item Updates are initially provisional
%     }
%     \uncover<2->{
%     \item Transactions distinguish between their first and second updates using labels
%     }
%     \uncover<3->{
%     \item When a transaction attempts to update a given record, it can identify all other transactions that have earlier updated that record provisionally
%     }
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Interference-free Update}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}
%       \uncover<1->{
%         \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
%         \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
%         \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
%         \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
%         \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
%       }
%       \uncover<2->{
%         \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
%       }
%       \uncover<3->{
%         \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
%       }
%       \uncover<3->{
%         \draw [red,->,>=stealth] (0,3.5) node [] {} -- (4,0.3) node (TX) [] {};
%       }
%       \uncover<4->{
%         \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
%       }
%     \end{tikzpicture}
%   \end{figure}

% \end{frame}

% \begin{frame}
%   \frametitle{Collision Detection}
%   \begin{center}
%     \textbf{Rule: For any ``1'' seen, there must be a ``2'' in the opposite end. Else, abort}
%   \end{center}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}
%       \uncover<2->{
%         \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
%         \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
%         \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
%         \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
%         \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
%       }
%       \uncover<2->{
%         \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
%       }
%       \uncover<3->{
%         \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
%       }
%       \uncover<4-11>{
%         \draw [red,->,>=stealth] (0,3.5) -- (2,1.9);
%       }
%       \uncover<5-10>{
%         \node [] at (-0.5,2.6) {\textcolor{blue}{$T_B$}};
%       }
%       \uncover<6-10>{
%         \node [] at (-0.5,2.1) {\textcolor{blue}{$1$}};
%       }
%       \uncover<7-7>{
%         \draw [blue,->,>=stealth] (0,2.1) -- (2,1.8);
%       }
%       \uncover<8-10>{
%         \draw [blue,->,>=stealth] (0,2.1) -- (4,1.5);
%         \node [] at (4.5,1.5) {\textcolor{blue}{$2$}};
%       }
%       \uncover<9-9>{
%         \node [] at (6,2) {$T_R$'s 2 is missing};
%       }
%       \uncover<10-10>{
%         \node[] at (6,2) {$T_B$ ABORTS};
%       }
%       \uncover<12->{
%         \draw [red,->,>=stealth] (0,3.5)  -- (4,0.3);
%       }
%       \uncover<13->{
%         \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
%       }
%       \uncover<14->{
%         \node[] at (6,2) {$T_R$ SUCCEEDS};
%         }
%     \end{tikzpicture}
%   \end{figure}
% \end{frame}

% %------------------------------------------------
% %Problem 2: Edge-Order Consistency
% % ------------------------------------------------

% \begin{frame}
%   \frametitle{Edge-Order Consistency}
%   A second consistency problem related to distributed edges:
%   \begin{itemize}
%     \uncover<2->{
%     \item Consider 2 transactions that update $2$ or more of the \textit{same} distributed edges
%     }
%     \uncover<3->{
%     \item Each update is reciprocally consistent
%     }
%     \uncover<4->{
%     \item Then if $T_R$ updates before $T_B$ on a given edge then this order should be preserved across all edges
%     }
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Edge-Order Consistency Violation}
%   Consider two edges,
%   \begin{figure}[h!]
%     \centering
%     \resizebox{1\textwidth}{!}{
%       \begin{tikzpicture}[node distance=5.2cm]
%         \node[] at (2.6,2.5) {\large{$E$}};
%         \node[] at (13,2.5) {\large{$E'$}};
%         \node (n1) [bignode] {\texttt{Flight}};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5.2cm,label=above:{$S_2$}] {};
%         \node (n2) [bignode, right of=n1] {\texttt{Seat}};
%         \draw [local] (n1) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n2);
%         \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textit{Seat Available In Flight}};
%         \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5.2cm,yshift=-2cm] {\textit{In Flight Seat Available}};
%          \draw [<->,dashed](q1) -- (q2);

%         \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
%         \node (n4) [bignode, right of=n3] {\texttt{Room}};
%         \draw [local] (n3) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n4);
%         \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10.4cm,yshift=-2cm] {\textit{Room Available In Hotel}};
%         \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15.6cm,yshift=-2cm] {\textit{In Hotel Room Available}};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15.6cm,label=above:{$S_4$}] {};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10.4cm,label=above:{$S_3$}] {};
%         \draw [<->,dashed](q3) -- (q4);

%       \end{tikzpicture}
%     }
%     \caption{Two distributed edges}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Edge-Order Consistency Violation}
%   Two concurrent transactions by a travel agent:
%   \begin{itemize}
%   \item Requests to book room and seat for Mr Red, $T_R$
%   \item Requests to book room and seat for Mr Blue, $T_B$
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Edge-Order Consistency Violation}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}
%       \uncover<1->{

%         \draw [black,->,>=stealth] (0,0) -- (10,0) node [label=below:{$time$}] {};
%         \draw [black,-,>=stealth] (2,4) node [label={[red]left:{$T_R$}}] {} -- (8,4);
%         \draw [black,-,>=stealth] (8,4.1) node [] {} -- (8,3.9);
%         \draw [black,-,>=stealth] (2,4.1) node [] {} -- (2,3.9);
%         \draw [black,-,>=stealth] (4,2) node [label={[blue]left:{$T_B$}}] {} -- (7,2);
%         \draw [black,-,>=stealth] (4,2.1) node [] {} -- (4,1.9);
%         \draw [black,-,>=stealth] (7,2.1) node [] {} -- (7,1.9);
%       }
%       \uncover<2->{
%         \draw [black,dashed] (2.5,0) node [label=below:{$t_1$}] {} -- (2.5,4) node [label=above:{$E$}] {};
%       }
%       \uncover<3->{
%         \draw [black,dashed] (4.5,0) node [label=below:{$t_2$}] {} -- (4.5,2) node [label=above:{$E$}] {};
%       }
%       \uncover<4->{

%         \draw [black,dashed] (6.5,0) node [label=below:{$t_3$}] {} -- (6.5,2) node [label=above:{$E'$}] {};
%       }
%       \uncover<5->{

%         \draw [black,dashed] (7.5,0) node [label=below:{$t_4$}] {} -- (7.5,4) node [label=above:{$E'$}] {};
%       }
%       \uncover<6->{
%         \draw (8.3,2) node [label=right:{$E : T_R \rightarrow T_B$}] {} ;
%         \draw (8.3,1.5) node [label=right:{$E' : T_B \rightarrow T_R$}] {} ;
%         \draw (8.3,1) rectangle (11,2.5);
%       }
%     \end{tikzpicture}
%     \caption{Edge-order consistency violation}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Edge-Order Consistency Violation}
%   \begin{figure}[h!]
%     \centering
%     \resizebox{1\textwidth}{!}{
%       \begin{tikzpicture}[node distance=5cm]
%         % edge labels
%         \node[] at (2.5,3) {\large{$E$}};
%         \node[] at (12.5,3) {\large{$E'$}};
%         % edge 1: flight -> seat
%         \node (n1) [bignode] {\texttt{Flight}};
%         \node (n2) [bignode, right of=n1] {\texttt{Seat}};
%         % edge 1: servers
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5cm,label=above:{$S_2$}] {};
%         % edge 1
%         \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);
%         % edge 1 information
%         \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
%         \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
%         \draw [<->,dashed](q1) -- (q2);

%         % edge 2: hotel -> room
%         \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
%         \node (n4) [bignode, right of=n3] {\texttt{Room}};
%         \draw [local] (n3) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n4);
%         \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
%         \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15cm,label=above:{$S_4$}] {};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10cm,label=above:{$S_3$}] {};
%         \draw [<->,dashed](q3) -- (q4);

%       \end{tikzpicture}
%     }
%     \caption{Edge-order consistency violation}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Our Solution: Order Arbitration}
%   \begin{itemize}
%   \uncover<1->{
%   \item Transactions collect predecessors from each update
%   }
%   \uncover<2->{
%   \item Centralized arbiter maintains some global state
%   }
%   \uncover<3->{
%   \item Global state used to detect edge-order violations
%   }
%   \uncover<4->{
%   \item Offending transactions are aborted
%   }
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Order Arbitration}
%     \begin{center}
%       \textbf{Rule: Collect predecessors after every successful update}
%     \end{center}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}
%       \uncover<1->{
%         \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
%         \draw [black,<-,>=stealth] (3,0) -- (3,4) ;
%         \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (3,4.5) node [label=above:{$S_2$}] {};
%         \draw [black,<-,>=stealth] (7,0) -- (7,4) ;
%         \draw [black,<-,>=stealth] (10,0) -- (10,4) ;
%         \draw [black,dashed,->,>=stealth] (7,4.5) node [label=above:{$S_3$}] {} -- node [midway,above] {$E'$} (10,4.5) node [label=above:{$S_4$}] {};
%       }
%       \uncover<2->{
%         \draw [blue,->,>=stealth] (0,3.2) node (TXS) [label=left:{\small{$T_B$}}] {} -- (3,3);
%         \draw [red,->,>=stealth] (3,3.6) node (TXS) [label=right:{\small{$T_R$}}] {} -- (0,3.4);
%         \draw (3.1, 2.8) node [label=right:{\small{\textcolor{black}{$T_B : \{{T_R}\}$}}}] {};
%         \draw (3.2,2.3) rectangle (5,3.2);
%       }
%       \uncover<3->{
%         \draw [blue,->,>=stealth] (7,2.3) node [label=left:{\small{$T_B$}}] {} -- (10,1.8);
%         \draw [red,->,>=stealth] (7,1.8) node [label=left:{\small{$T_R$}}] {} -- (10,1.2) ;
%         \draw (10.1, 0.8) node (TXS) [yshift=-1cm,label=right:{\small{\textcolor{black}{$T_R : \{{T_B}\}$}}}] {};
%         \draw (10.2,-0.6) rectangle (12,0.3);
%       }
%       \uncover<4->{
%         \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
%         (10.2,1.8) node[above,xshift=1cm, yshift=0.3cm] {ARBITER} -- (11,1.8);
%         \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
%         (10.2,1.2)   -- (11,1.2);
%       }
%     \end{tikzpicture}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Order Arbitration}
%       \begin{center}
%       \textbf{Rule: If transaction exists in hit list, abort. Else, merge predecessors into hit list}
%     \end{center}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}
%       \uncover<1->{ % 1 arbiter
%         \node (s1) [server,label=$Arbiter$] {};
%       }
%       \uncover<2->{ % 2 hit list
%         \node (a) [rectangle, label=above:{$Hit \ List$}, minimum width=2cm,minimum height=1cm,draw=black,xshift=0cm,yshift=0cm] {$T_z$ ,$T_a$ {\uncover<7->{,$T_R$}}};
%       } % uncover T_R when T_B processed
%       \uncover<3->{ % 3 queue
%               \node (q) [rectangle, label=above:{$Queue$},left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=0cm] {{\uncover<6->{$...$}}};
%               \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=0cm] {{\uncover<5-8>{$T_R$}}};
%         \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=0cm] {{\uncover<4-5>{$T_B$}}};
%         \node (q1) [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=-1cm] {{\uncover<6->{$...$}}};
%         \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=-1cm] {{\uncover<5-8>{$T_B$}}};
%         \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=-1cm] {{\uncover<4-5>{$T_R$}}};
%         \node [left of=q,xshift=0.2cm,label=left:{\small{\textcolor{black}{$tnx.id$}}}] {};
%         \node [left of=q1,xshift=0.2cm,label=left:{\small{\textcolor{black}{$pred$}}}] {};
%       }
%       \uncover<4-4> {
%         \node[right=0.5cm of s1] {$T_B$ \textbf{ENTERS QUEUE}};
%       }
%       \uncover<5-5> {
%         \node[right=0.5cm of s1] {$T_R$ \textbf{ENTERS QUEUE}};
%       }
%       \uncover<6-8> {
%         \node[right=0.5cm of s1,yshift=1cm] {$T_B$ \textbf{PROCESSED}};
%       }
%       \uncover<7-8> {
%         \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{ADDED TO HIT LIST}};
%       }
%       \uncover<8-8> {
%         \node[right=0.5cm of s1,,yshift=-1cm] {$T_B$ \textbf{SUCCEEDS}};
%       }
%       \uncover<9-> {
%         \node[right=0.5cm of s1,yshift=1cm] {$T_R$ \textbf{PROCESSED}};
%       }
%       \uncover<10-> {
%         \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{IN HIT LIST}};
%       }
%       \uncover<11-> {
%         \node[right=0.5cm of s1,yshift=-1cm] {$T_R$ \textbf{ABORTS}};
%       }
%     \end{tikzpicture}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Order Arbitration Protocol}
%   If a transaction updates $1$ or more distributed edges:
%   \begin{itemize}
%     \uncover<1->{
%     \item Collect all predecessors
%     }
%     \uncover<2->{
%     \item Go to arbiter
%     }
%     \uncover<3->{
%     \item If not in hit list then proceed and enter predecessors into the hit list
%     }
%     \uncover<4->{
%     \item Else in hit list then abort
%     }
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Edge Concurrency Control Protocol Summary}
%     \begin{enumerate}
%     \item Collision detection: enforces \textbf{reciprocal consistency} for distributed edges
%     \item Order arbitration: enforces \textbf{edge-order consistency} between transactions
%     \end{enumerate}
% \end{frame}

% %------------------------------------------------
% % Performance Evaluation
% % ------------------------------------------------

% \begin{frame}
%   \frametitle{Performance Evaluation}
%   Performance measures of interest:
%   \begin{itemize}
%   \item Average number of transactions that are aborted
%   \item Load at the arbiter
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Evaluation Approach}
%   \begin{enumerate}
%   \item Developed approximate model
%   \item Measure accuracy of model through simulation
%   \end{enumerate}
% \end{frame}

% \begin{frame}
%   \frametitle{Parameters}
%   \begin{itemize}
%   \item Database size
%   \item Transaction arrival rate
%   \item Average updates per transaction
%   \item Average arbiter service time
%   \item Average network delay
%   \end{itemize}
% \end{frame}

% %------------------------------------------------
% % Numerical and Simulation Results
% % ------------------------------------------------

% \begin{frame}
%   \frametitle{Aborts per second $(\boldsymbol{R})$ vs Database Size $(\boldsymbol{N})$}
%   \begin{figure}[h!]
%     \centering
%     \includegraphics[scale=0.35]{images/varying_num_edges.png}
%     \caption{TPS $= 1000$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
%   \begin{center}
%     Arbiter queue unstable at $\sim1500$ TPS
%   \end{center}
%   \begin{figure}[h!]
%     \centering
%     \includegraphics[scale=0.35]{images/varying_tnx_arrival_rate}
%     \caption{\textcolor{red}{size $= 10K$}, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
%   \begin{center}
%     Arbiter queue unstable at $\sim 1100$ TPS
%   \end{center}
%   \begin{figure}[h!]
%     \centering
%     \includegraphics[scale=0.35]{images/larger_network_delay.png}
%     \caption{size $= 10K$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, \textcolor{red}{av network delay $=10ms$}}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
%   \begin{center}
%     Arbiter queue unstable at $\sim 550$ TPS
%   \end{center}
%     \begin{figure}[h!]
%     \centering
%     \includegraphics[scale=0.35]{images/varying_updates_per_tnx.png}
%         \caption{size $= 10K$, \textcolor{red}{av updates/tnx $= 10$}, av arbiter service time $= 10ms$, av network delay $=5ms$}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Summary and Future Work}
%   \begin{itemize}
%     \uncover<1->{
%     \item Model accuracy similar to simulations under a variety of parameter settings
%     }
%     \uncover<2->{
%     \item Between 1-4\% transactions are aborted
%     }
%     \uncover<3->{
%     \item Improve accuracy of simulation
%     }
%   \end{itemize}
% \end{frame}

% \begin{frame}[standout]
%  Thanks for listening! \\
%   \vspace{4mm}
%   Any Questions? \\
%   \vspace{4mm}
%   {\normalsize Email: \href{mailto:j.waudby2@newcastle.ac.uk}{j.waudby2@newcastle.ac.uk}} \\
%   {\normalsize Twitter: \href{https://twitter.com/waudberry_7}{@waudberry\_7}}
% \end{frame}

\end{document}